---
title: "Experiment 0 analysis"
output: html_document
---

# libraries

```{r}
library(rjson)
library(purrr)
library(lme4)
library(lmerTest)
library(boot)
library(tidyverse)
```

```{r}
theme_set(theme_classic())
```

# Load data

```{r data loading}
data_dir = "../anonymized_data/ex0"
data_files = list.files(data_dir)

df_graph_trials = data.frame()

graph_trials_to_df = function(list_data) {
  df = tibble::as_tibble(list_data[c("trial_type", "trial_index", "graph_trial_type", "trajectory", "keypresses", "keypress_rts", "keyups", "keyup_rts")])
  df$nodes_to_keys = rep(list(list_data$nodes_to_keys), nrow(df))
  if ("graph_name" %in% names(list_data)) {
    df$graph_name = list_data$graph_name
  } else {
    df$graph_name = NA
  }
  return(df)
}

for (filename in data_files) {
  list.this_subject_data = fromJSON(file=sprintf("%s/%s", data_dir, filename)) 
  df_this_subject_data = graph_trials_to_df(list.this_subject_data[[5]]) %>%
    bind_rows(graph_trials_to_df(list.this_subject_data[[7]]))
  df_this_subject_data$subject_id = str_split(filename, '\\.', simplify=T)[1, 1] 
  
  df_graph_trials = bind_rows(df_graph_trials, df_this_subject_data)
  
}
```

## Load graph structures

```{r graph loading}
three_rooms_adjacency = as.matrix(read_csv('auxiliary_data/tr_adjacency.csv', col_names=F))
fixed_random_adjacency = as.matrix(read_csv('auxiliary_data/rand_adjacency.csv', col_names=F))
```

# data wrangling
fix for missing condition info in some participants
```{r}
# find subjects with missing info
df_graph_trials %>% 
  filter(is.na(graph_name)) %>% 
  pull(subject_id) %>% 
  unique()

lookup_graph_name = function(subject_id, graph_trial_type) {
  condition = case_when(subject_id == "1" ~ "b",
                        subject_id == "2" ~ "c",
                        subject_id == "3" ~ "c",
                        subject_id == "4" ~ "b",
                        subject_id == "5" ~ "a",
                        subject_id == "6" ~ "b",
                        subject_id == "7" ~ "d",
                        subject_id == "8" ~ "a",
                        subject_id == "9" ~ "d",
                        subject_id == "10" ~ "b",
                        subject_id == "11" ~ "c",
                        subject_id == "12" ~ "d",
                        subject_id == "13" ~ "d",
                        subject_id == "14" ~ "d",
                        subject_id == "15" ~ "a",
                        subject_id == "16" ~ "c",
                        subject_id == "17" ~ "c",
                        subject_id == "18" ~ "a",
                        subject_id == "19" ~ "a",
                        subject_id == "20" ~ "c",
                        subject_id == "21" ~ "a",
                        subject_id == "22" ~ "c",
                        subject_id == "23" ~ "d",
                        subject_id == "24" ~ "c",
                        subject_id == "25" ~ "d",
                        subject_id == "26" ~ "b",
                        subject_id == "27" ~ "b",
                        subject_id == "28" ~ "d",
                        subject_id == "29" ~ "a",
                        subject_id == "30" ~ "c",
                        subject_id == "31" ~ "a",
                        subject_id == "32" ~ "a",
                        subject_id == "33" ~ "d",
                        subject_id == "34" ~ "b",
                        subject_id == "35" ~ "d",
                        subject_id == "36" ~ "b",
                        subject_id == "37" ~ "b",
                        subject_id == "38" ~ "b",
                        subject_id == "39" ~ "a",
                        subject_id == "40" ~ "c",
                        T ~ "NA")
  
  name = case_when(condition == "a" & graph_trial_type == "key_combination" ~ "three_rooms", 
                   condition == "a" & graph_trial_type == "letter" ~ "fixed_random_graph_0",  
                   condition == "b" & graph_trial_type == "key_combination" ~ "three_rooms", 
                   condition == "b" & graph_trial_type == "letter" ~ "three_rooms",  
                   condition == "c" & graph_trial_type == "key_combination" ~ "fixed_random_graph_0", 
                   condition == "c" & graph_trial_type == "letter" ~ "fixed_random_graph_0",  
                   condition == "d" & graph_trial_type == "key_combination" ~ "fixed_random_graph_0", 
                   condition == "d" & graph_trial_type == "letter" ~ "three_rooms",  
                   T ~ "NA")
  
  return(name)
}

df_graph_trials = df_graph_trials %>%
  mutate(graph_name = lookup_graph_name(subject_id, graph_trial_type))

any(df_graph_trials$graph_name == "NA" | is.na(df_graph_trials$graph_name))
xtabs(~ graph_trial_type + graph_name + condition, df_graph_trials)
```

```{r}
df_graph_trials = df_graph_trials %>%
  group_by(subject_id, graph_trial_type) %>%
  mutate(time=1:n()) %>%
  ungroup()
```

```{r isomorphic?}
df_graph_trials = df_graph_trials %>%
  group_by(subject_id) %>%
  mutate(condition = ifelse(length(unique(graph_name)) == 1,
                            "isomorphic",
                            "nonisomorphic")) %>%
  ungroup()
```


```{r marking correct}
df_graph_trials = df_graph_trials %>%
  rowwise() %>%
  mutate(target = ifelse(graph_trial_type == 'letter', 
                         list(nodes_to_keys[[trajectory+1]]),
                         list(nodes_to_keys[[trajectory+1]]))) %>%
  ungroup() %>%
  mutate(correct = map_int(keypresses, length) == map_int(target, length))
```

```{r transition random or within graph}
df_graph_trials = df_graph_trials %>%
  rename(current_node=trajectory) %>%
  mutate(current_node = current_node + 1) %>% # easier for 1-based indexing in R
  group_by(subject_id, graph_trial_type) %>%
  mutate(previous_node = lag(current_node, 1)) %>%
  rowwise() %>%
   mutate(transition_in_graph = case_when(graph_name == "three_rooms" ~ three_rooms_adjacency[previous_node, current_node] == 1,
                                         graph_name == "fixed_random_graph_0" ~ fixed_random_adjacency[previous_node, current_node] == 1,
                                         T ~ NA)) %>%
  ungroup()
```

there are complex issues about what is a first RT on incorrect key-combination trials -- should we interpret the first key-press as part of a two-key combination or not?

```{r first RT}
df_graph_trials = df_graph_trials %>%
  rowwise() %>%
  mutate(first_response_RT = case_when(trial_type == "letter" ~ keypress_rts[1],
                              length(target) == 1 ~ keypress_rts[1],
                              length(target) == 2 ~ keypress_rts[2],
                              T ~ as.double(NA))) %>%
  ungroup()
```

A few other things that may affect RT: specific key-presses + whether they are combos

```{r}
df_graph_trials = df_graph_trials %>% 
  mutate(target_str = map_chr(target, function(x) {paste(x, collapse="")}),
         target_combo = map_lgl(target, function(x) {length(x) > 1}))
```

# basic RT differences

```{r}
df_subj_rt_summaries = df_graph_trials %>%
  filter(first_response_RT < 2500) %>%
  group_by(subject_id, condition, graph_name, graph_trial_type, transition_in_graph, correct) %>%
  summarize(mean_fRT = mean(first_response_RT),
            sd_fRT = sd(first_response_RT)) %>%
  ungroup() 
  
```

## learning effect?
```{r}
df_subj_rt_summaries %>%
  group_by(graph_trial_type, transition_in_graph, correct) %>%
  summarize(mean_mean_fRT = mean(mean_fRT),
            sd_mean_fRT = sd(mean_fRT)) %>%
  ungroup()
```

```{r}
ggplot(df_subj_rt_summaries %>%
         filter(!is.na(transition_in_graph)), 
       aes(x=transition_in_graph, y=mean_fRT, color=transition_in_graph)) +
  geom_point(stat="summary",
           fun.y="mean") +
  geom_errorbar(stat="summary",
                fun.data="mean_cl_normal",
                width=0.25) +
   facet_grid(correct ~ graph_trial_type, scales="free_y")
ggsave("results/plots/basic_learning_effect_RT_raw.png", width=4, height=6)
```

## difference scores

```{r}
df_subj_rt_diffs = df_subj_rt_summaries %>%
  select(-sd_fRT) %>%
  filter(!is.na(transition_in_graph)) %>%
  mutate(transition_in_graph = ifelse(transition_in_graph, "in_graph", "out_of_graph")) %>%
  spread(transition_in_graph, mean_fRT) %>%
  mutate(rt_diff = out_of_graph - in_graph)
```

```{r}
ggplot(df_subj_rt_diffs %>%
         filter(correct),
       aes(x=graph_name, y=rt_diff, fill=graph_trial_type)) +
  geom_bar(stat="summary",
           fun.y="mean",
           position="dodge") +
  geom_errorbar(stat="summary",
                fun.data="mean_cl_boot",
                position=position_dodge(width=0.9),
                width=0.25)

ggsave("results/plots/basic_learning_effect_RT.png", width=4, height=6)
```

## isomorphic mean effect on RT?


```{r isomorphic mean effect on RT (correct only)}
ggplot(df_subj_rt_summaries %>%
         filter(!is.na(transition_in_graph),
                correct), 
       aes(x=condition, y=mean_fRT, fill=condition)) +
  geom_bar(stat="summary",
           fun.y="mean") +
  geom_errorbar(stat="summary",
                fun.data="mean_cl_normal") +
  facet_grid(transition_in_graph ~ graph_trial_type)


ggsave("results/plots/isomorphic_effect_RT_raw.png", width=4, height=6)
```

## isomorphic difference

```{r}
ggplot(df_subj_rt_diffs %>%
         filter(correct),
       aes(x=graph_name, y=rt_diff, fill=condition)) +
  geom_bar(stat="summary",
           fun.y="mean",
           position="dodge") +
  geom_errorbar(stat="summary",
                fun.data="mean_cl_boot",
                position=position_dodge(width=0.9),
                width=0.25) +
  facet_grid(~ graph_trial_type)

ggsave("results/plots/transfer_effect_RT_diff.png", width=4, height=6)
```


# effects on % correct on first try

## isomorphic mean effect

```{r}
df_subj_corr_summaries = df_graph_trials %>%
  group_by(subject_id, condition, graph_name, graph_trial_type, transition_in_graph) %>%
  summarize(mean_correct = mean(correct),
            sd_correct = sd(correct)) %>%
  ungroup() 
  
```

```{r}
df_subj_corr_summaries %>%
  group_by(graph_trial_type, transition_in_graph) %>%
  summarize(mean_mean_correct = mean(mean_correct),
            sd_mean_correct = sd(mean_correct)) %>%
  ungroup()
```

```{r}
df_subj_corr_summaries %>%
  group_by(condition, graph_trial_type, transition_in_graph) %>%
  summarize(mean_mean_correct = mean(mean_correct),
            sd_mean_correct = sd(mean_correct)) %>%
  ungroup() %>%
  filter(!is.na(transition_in_graph)) %>%
  arrange(transition_in_graph, graph_trial_type, condition)

```

```{r}
ggplot(df_subj_corr_summaries %>%
         filter(!is.na(transition_in_graph)), 
       aes(x=transition_in_graph, y=mean_correct, color=transition_in_graph)) +
  geom_point(stat="summary",
             fun.y="mean") +
  geom_errorbar(stat="summary",
                fun.data="mean_cl_boot") +
  facet_grid(~ graph_trial_type) 

ggsave("results/plots/basic_learning_effect_pct_correct.png", width=4, height=6)
```

```{r}
ggplot(df_subj_corr_summaries %>%
         filter(!is.na(transition_in_graph),
                transition_in_graph), 
       aes(x=condition, y=mean_correct, color=condition)) +
  geom_point(stat="summary",
             fun.y="mean") +
  geom_errorbar(stat="summary",
                fun.data="mean_cl_boot") +
  facet_grid(graph_name ~ graph_trial_type) 

ggsave("results/plots/transfer_effect_pct_correct.png", width=4, height=6)
```

## differences

```{r}
df_subj_corr_diffs = df_subj_corr_summaries %>%
  select(-sd_correct) %>%
  filter(!is.na(transition_in_graph)) %>%
  mutate(transition_in_graph = ifelse(transition_in_graph, "in_graph", "out_of_graph")) %>%
  spread(transition_in_graph, mean_correct) %>%
  mutate(corr_diff = out_of_graph - in_graph)
```

```{r}
ggplot(df_subj_corr_diffs,
       aes(x=graph_name, y=corr_diff, fill=condition)) +
  geom_bar(stat="summary",
           fun.y="mean",
           position="dodge") +
  geom_errorbar(stat="summary",
                fun.data="mean_cl_boot",
                position=position_dodge(width=0.9),
                width=0.25) +
  facet_grid(~ graph_trial_type)

ggsave("results/plots/isomorphic_effect_corr.png", width=4, height=6)
```

# correct model 

```{r}
corr_out_model = glmer(correct ~ condition * graph_trial_type + graph_name  + target_combo + poly(time, 2) + (1 | subject_id) + (1 | target_str), 
                   family="binomial",
                   data=df_graph_trials %>%
                     filter(!is.na(transition_in_graph),
                            !transition_in_graph))
summary(corr_out_model)
```

```{r}
corr_in_model = glmer(correct ~ condition * graph_trial_type + graph_name  + target_combo + poly(time, 2) + (1 | subject_id) + (1 | target_str), 
                   family="binomial",
                   data=df_graph_trials %>%
                     filter(!is.na(transition_in_graph),
                            transition_in_graph))
summary(corr_in_model)
```


```{r}
corr_full_model = glmer(correct ~ condition * graph_trial_type * transition_in_graph + graph_name  + target_combo + poly(time, 2) + (1 | subject_id) + (1 | target_str), 
                   family="binomial",
                   data=df_graph_trials %>%
                     filter(!is.na(transition_in_graph)))
summary(corr_full_model)
```

# rt proper analysis
```{r}
ggplot(df_graph_trials,
       aes(x=first_response_RT)) +
  geom_histogram() +
  xlim(0, 5000)
```

```{r}
ggplot(df_graph_trials %>% 
         filter(correct,
                first_response_RT < 5000),
       aes(x=time, color=condition, linetype=transition_in_graph, y=first_response_RT)) +
  geom_line(stat="summary",
            fun.y="mean") +
  facet_grid(graph_name ~ graph_trial_type)
```


```{r}
rt_model = lmer(first_response_RT ~ condition * graph_trial_type + graph_name + transition_in_graph + target_combo + poly(time, 2) + (1 + poly(time, 2) | subject_id) + (1 | target_str), 
                data=df_graph_trials %>%
                  filter(correct, 
                         !is.na(transition_in_graph),
                         first_response_RT < 5000))
summary(rt_model)
```

